---
title: "deal.Rmd"
author: "Ethan Decker"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Your task consists of arguing empirically whether Marilyn Vos Savant is 
correct. In an R script named `deal.R` (or `deal.Rmd`), write a simulation 
study that replicates the game show as much as possible (i.e., randomly 
choosing doors, etc.). Implement both strategies (i.e., "do not switch" versus 
"switch").  Give the probability of winning under each of the strategies. 
Quantify your uncertainity about the emperically estimated probabilities by 
forming a confidence intervals on the proportions of wins.
Let's set up one round of a game:
```{r}
#we need three doors to choose from
doors <- c(1,2,3)
#we need to pick one of three doors at random
my.door <- sample(doors, size = 1, prob = c(1/3,1/3,1/3))
#we need the other two doors...
other.doors <- doors[doors != my.door] #success! we have 3 doors with one picked at random!
```
Now that we have doors we really only need one prize randomly hidden behind one of the doors...
```{r}
Big.prize <- sample(doors, size = 1, prob =  c(1/3,1/3,1/3))
#we need to open up one of the other doors and throw it away
swap.door <- sample(other.doors[other.doors == Big.prize], size = 1)
```
Now we can set up a monte carlo simulation for a strategy, Let's not swap doors. 
1st Simulation, Strategy One don't switch your door
```{r}
#write a function that creates this simulation
doors <- c(1,2,3)
my.door <- sample(doors, size = 1, prob = c(1/3,1/3,1/3))
other.doors <- doors[doors != my.door]
Big.prize <- sample(doors, size = 1, prob =  c(1/3,1/3,1/3))
swap.door <- doors[doors != my.door]
strat.1 <- function() {
  doors <- c(1,2,3)
  my.door <- sample(doors, size = 1, prob = c(1/3,1/3,1/3))
  other.doors <- doors[doors != my.door]
  Big.prize <- sample(doors, size = 1, prob =  c(1/3,1/3,1/3))
  y <- 1
  x <- 0
  if(my.door == Big.prize)
  return(y)
    else(my.door != Big.prize)
  return(x)
}
```
The first strategy is just a 1/3 chance you pick the right door. So the equation is simply if you picked the right door or not. The rest other door opened and the option to swap don't really matter in this case if the prize is assigned before you pick your door and your strategy is to not swap doors. 
The monte Carlo simulation would be this then:
```{r}
sim <- function(){
  replicate(1000, strat.1())
}
mc_est <- replicate(1000, sim())
mc_mean <- mean(mc_est)
mc_sd <- sd(mc_est)
x_bar <- mc_mean
sd_x <- sd(mc_est)
sqrt_n <- sqrt(length(mc_est))
ci <- x_bar + c(-1,1)*qnorm(1-.025)*sd_x/sqrt_n
cat("Confidence Interval for the mean of Strategy 1 =", ci, "\n")
cat("Estimated Chance of winning with Strategy 1 =", x_bar, "\n")
```
Now lets work on strategy 2, switching your door
```{r}
#lets start with some doors, I like doors
doors <- c(1,2,3)
Big.prize <- sample(doors, size = 1)
my.door <- sample(doors, size = 1)
other.doors <- doors[doors != my.door]
trash.door <- other.doors[other.doors != Big.prize]
remaining.doors <- as.numeric(sample(trash.door, size = 1))
swap.door <- as.numeric(sample(remaining.doors, size = 1))
#okay now lets call out a function that uses this information
strat.2 <- function(x){
  doors <- 1:3
  Big.prize <- sample(1:3, size = 1, prob = c(1/3,1/3,1/3))
  my.door <- sample(1:3, size = 1, prob = c(1/3,1/3,1/3))
  other.doors <- doors[doors != my.door]
  trash.door <- other.doors[other.doors != Big.prize]
  if(length(trash.door) == 2)
    remaining.doors <- sample(trash.door, size = 1, prob = c(.5, .5))
  else(length(trash.door) == 1)
  remaining.doors <- sample(trash.door, size = 1, prob = NULL)
  final.door <- other.doors[other.doors != remaining.doors]
  swap.door <- as.numeric(sample(final.door, size = 1))
  x <- 1
  y <- 0
  if(my.door == Big.prize)
    return(y)
  else(swap.door == Big.prize)
    return(x)
}
strat.2()
#now lets simulate it a few times
sim.2 <- function(){
  replicate(1000, strat.2())
}
mc_est.2 <- replicate(100, sim.2())
mc_mean.2 <- mean(mc_est.2)
x_bar <- mc_mean.2
sd_x <- sd(mc_est.2)
sqrt_n <- sqrt(length(mc_est.2))
ci <- x_bar + c(-1,1)*qnorm(1-.025)*sd_x/sqrt_n
cat("Confidence Interval for the mean of Strategy 1 =", ci, "\n")
cat("Estimated Chance of winning with Strategy 2 =", x_bar, "\n")
```



