---
title: "SunSpot.Rmd"
author: "Ethan Decker"
date: "4/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Read in data:
```{r}
sunspots <- as.vector(sunspot.month)

library(ggplot2)

ggplot(mapping = aes(x = seq_along(sunspots), y = sunspots)) +
  geom_line() +
  theme_bw() +
  labs(x = "Days",
       y = "Sunspots",
       title = "Sunspots over time")
```

1: Calculate the observed auto-correlation (10 points)
```{r}
#observed coefficient is
observed_cor <- cor(sunspots[-length(sunspots)], sunspots[-1])
observed_cor[1]
```
2: Perform a permutation test with 10,000 samples and store the results (10 points)
```{r}
n_samples <- 10000
permute_results <- replicate(n_samples, {
  new_spots <- sample(sunspots)
  x <- cor(new_spots[-length(new_spots)], new_spots[-1])
  x[1]
})
```
3:Plot the density of the results from step 3 along with a vertical red line for the observed auto-correlation calculated in step 2 (10 points)
```{r}
plot(density(permute_results), xlim = c(observed_cor, max(permute_results)))
abline(v = observed_cor, col = "red")
```
4:Using the results from 2 and 3, calculate an estimated P value for the observed auto-correlation (5 points)
```{r}
p_value <- mean(abs(permute_results) >= abs(observed_cor[1]))
```
5: Calculate a 95% confidence interval on the P value you estimated in step 5 (5 points)
```{r}
ci <- p_value + c(-1, 1) * qnorm(.975) * sqrt(p_value * (1 - p_value) / n_samples)
c(lower = ci[1],
  p_value = p_value,
  upper = ci[2])
```
6:Interpret your results
The odds of getting sample data as extreme or more extreme than the data given is basically 0. That means we have sufficient evidence to reject the null hypothesis. There is a positive autocorrelation between for sunspots. 
