---
title: "Tp Problem"
author: "Ethan Decker"
date: "3/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 you have 8, 284 sheet rolls left. You estimate the following:

The number of times you need toilet paper each day is either 1, 2, or 3, with equal likelihood for each
Each time you need toilet paper, the number of squares used is defined by a Poisson distribution with λ=10
1.) Define a function called tp that uses the above information and returns the number of days until you run out of TP
```{r}
tp <- function(rolls, sheets){
  sheets <- rolls*sheets
  x <- 0
  while(sheets > 0) {
  total.days <- x
  #gives a value for number of times a day you need toilet paper with equal likelihood
  times.per.day <- sample(c(1,2,3), 1, prob = c(1/3,1/3,1/3))
  #this gives how many sheets of tp you use that day
  used.tp <- sum(rpois(times.per.day, lambda = 10))
  #the awaited hour is the awaited hour where you are out of toilet paper
  awaited.hour <- sheets-used.tp
  #Now I write over the variable sheets to restart the loop
  sheets <- awaited.hour
  #next line counts how many times this while loop runs
  x <- x+1
  }
  return(x)
}
tp(rolls = 8, sheets = 284)
```
2.2)  Using the function defined in 1, run this simulation 10,000 times and store the results.
```{r}
tp.sim <- replicate(10000, tp(rolls = 8, sheets = 284))
```
2.3) Estimate the number of days you have tp
```{r}
Est.days <- mean(tp.sim)
Est.days
```
2.4) 95% confidence interval to go with estimate from 2.3
```{r}
Confidence.Interval <- function(x){
  x_bar <- mean(x)
sd_x <- sd(x)
sqrt_n <- sqrt(length(x))
alpha <- 0.05

ci <- x_bar + c(-1, 1) * qnorm(1 - alpha/2) * sd_x / sqrt_n
return(ci)
}
Interval <- Confidence.Interval(tp.sim)
cat("Confidence Interval for the Awaited No TP Day =", Interval, "\n")
```

